### 2023-01-27 Buffer_CBD.qmd
Commit.

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.28.2-Firenze">
  <identifier></identifier>
  <parentidentifier></parentidentifier>
  <language></language>
  <type>dataset</type>
  <title></title>
  <abstract></abstract>
  <links/>
  <fees></fees>
  <encoding></encoding>
  <crs>
    <spatialrefsys nativeFormat="Wkt">
      <wkt>PROJCRS["ETRS89 / UTM zone 33N",BASEGEOGCRS["ETRS89",ENSEMBLE["European Terrestrial Reference System 1989 ensemble",MEMBER["European Terrestrial Reference Frame 1989"],MEMBER["European Terrestrial Reference Frame 1990"],MEMBER["European Terrestrial Reference Frame 1991"],MEMBER["European Terrestrial Reference Frame 1992"],MEMBER["European Terrestrial Reference Frame 1993"],MEMBER["European Terrestrial Reference Frame 1994"],MEMBER["European Terrestrial Reference Frame 1996"],MEMBER["European Terrestrial Reference Frame 1997"],MEMBER["European Terrestrial Reference Frame 2000"],MEMBER["European Terrestrial Reference Frame 2005"],MEMBER["European Terrestrial Reference Frame 2014"],ELLIPSOID["GRS 1980",6378137,298.257222101,LENGTHUNIT["metre",1]],ENSEMBLEACCURACY[0.1]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],ID["EPSG",4258]],CONVERSION["UTM zone 33N",METHOD["Transverse Mercator",ID["EPSG",9807]],PARAMETER["Latitude of natural origin",0,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8801]],PARAMETER["Longitude of natural origin",15,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8802]],PARAMETER["Scale factor at natural origin",0.9996,SCALEUNIT["unity",1],ID["EPSG",8805]],PARAMETER["False easting",500000,LENGTHUNIT["metre",1],ID["EPSG",8806]],PARAMETER["False northing",0,LENGTHUNIT["metre",1],ID["EPSG",8807]]],CS[Cartesian,2],AXIS["(E)",east,ORDER[1],LENGTHUNIT["metre",1]],AXIS["(N)",north,ORDER[2],LENGTHUNIT["metre",1]],USAGE[SCOPE["Engineering survey, topographic mapping."],AREA["Europe between 12°E and 18°E: Austria; Denmark - offshore and offshore; Germany - onshore and offshore; Norway including Svalbard - onshore and offshore."],BBOX[46.4,12,84.42,18]],ID["EPSG",25833]]</wkt>
      <proj4>+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs</proj4>
      <srsid>2106</srsid>
      <srid>25833</srid>
      <authid>EPSG:25833</authid>
      <description>ETRS89 / UTM zone 33N</description>
      <projectionacronym>utm</projectionacronym>
      <ellipsoidacronym>EPSG:7019</ellipsoidacronym>
      <geographicflag>false</geographicflag>
    </spatialrefsys>
  </crs>
  <extent/>
</qgis>