### 2023-02-19 MSB106_Assignment.qmd
.

---
title: "MSB106 - Assignment"
author: "Sindre H. Øveraas, Alen Colakovic, Mona Lisa Jones & Sebastian M. Fløysand"
format:
 pdf:
  fontsize: "10"
  documentclass: article
  margin-left: 2cm
  margin-right: 2cm
papersize: A4
editor: visual
---

```{r, message=FALSE, echo=FALSE}
library(rgdal)
library(dplyr)
library(RSQLite)
library(sf)
library(tidyverse)
library(readr)
library(ggplot2)
library(kableExtra)
library(huxtable)
```

```{r, message=FALSE, echo=FALSE}
NOR_CBD <- read_csv("NOR_CBD.csv")
Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
Zonal_CBD <- read_csv("Zonal_CBD.csv")
Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv") |>
  rename("HubName" = fid)
Zonal_Den <- read.csv("Zonal_Den.csv") |>
  select(fid, X_sum)
Dist_Comp_Den <- read.csv("Dist_Comp_Den.csv") |>
  select(fid, distance)
NOR_KOMM <- read_csv("NOR_KOMM.csv")
NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
```

```{r, echo=FALSE}
Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
  rename("knr" = kommunenummer, "DistMal" = HubDist, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
```

```{r, echo=FALSE}
Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
  select(knr, DistMal, HubName, Turnover_capita_retail, shop, Vinmonopolet) |>
  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
```

```{r, echo=FALSE, warning=FALSE}
Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
Commune_Data$Winemonopoly <- as.integer(as.logical(Commune_Data$Winemonopoly))
```

```{r, echo=FALSE}
Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubName" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
```

```{r, echo=FALSE}
Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
```

```{r, echo=FALSE}
Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
  rename("DistComp" = distance)
```

```{r, echo=FALSE}
Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
  select(fid, Juridisk.n, Antall.ans, Driftsres, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
```

```{r, echo=FALSE, warning=FALSE}
Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
Dentist_Data$Winemonopoly <- as.integer(as.logical(Dentist_Data$Winemonopoly))
```

```{r, echo=FALSE}
kable(summary(Commune_Data))
```

```{r, echo=FALSE}
kable(summary(Dentist_Data[3:6]))
```

```{r, echo=FALSE}
kable(summary(Dentist_Data[7:10]))
```

```{r, echo=FALSE}
kable(summary(Dentist_Data[11:12]))
```

```{r, echo=FALSE}
#| label: fig-1
#| tbl-cap: "Capita in retail compared to mall distance"
Commune_Data |>
  ggplot(aes(x = SizeMall, y = Turnover_capita_retail, colour = SizeMall)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
  theme_bw()
```

```{r,echo=FALSE}
#| label: fig-2
#| tbl-cap: "Dentist income compared to CBD distance"
Dentist_Data |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r,echo=FALSE}
Dentist_Data_LIMIT <- Dentist_Data |>
  filter(Sum.salgsi < 100000000, DistCBD < 250,
         Driftsres < 20000000, Driftsres > -5000000, DistMal < 80)
```

```{r, echo=FALSE}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r, echo=FALSE}
Dentist_Data |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r, echo=FALSE}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r, echo=FALSE}
Dentist_Data_LIMIT |>
  ggplot(aes(x = X_sum, y = Sum.salgsi, colour = Sum.salgsi)) +
  geom_point(lwd = .9) +
  labs(x = "Pop density", y = "Income") +
  theme_bw()
```

```{r, echo=FALSE}
lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
```

```{r, echo=FALSE}
huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
```

```{r, echo=FALSE}
Den_W_Wine <- Dentist_Data_LIMIT |>
  filter(Winemonopoly == 1)
```

```{r, echo=FALSE}
Den_WO_Wine <- Dentist_Data_LIMIT |>
  filter(Winemonopoly == 0)
```

```{r, echo=FALSE}
lm2 <- lm(Turnover_capita_retail ~ DistMal + SizeMall + (DistMal*SizeMall) + Winemonopoly, data = Commune_Data)
```

```{r, echo=FALSE}
huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
```

## Econometric approach